\mytitle

### 2016-08-28 pdocs/thesis/article-section-2.tex
added more details about maximum forces in article section 2

% pdf page 33, notation in typical ODEs
\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$\vec{r}_i$ & position of center of mass for body i  \\ \hline
$\vec{q}_i$ & orientation as a quaternion for body i  (\ref{eq:vectorqi})  \\ \hline
$\vec{p}_i$ & contact or joint point i  \\ \hline
$\vec{r}_{ki}$ & $\vec{p}_k - \vec{r}_i$  \\ \hline
$\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
$Q_i$ & rotation matrix $Q_i \in  \mathbb{R}^{4 \times 3}$ (\ref{eq:matrixQi})  \\ \hline
$S$ & generalized transformation matrix $ S \in \mathbb{R}^{7n \times 6n}$ (\ref{eq:matrixS}) \\ \hline
$\vec{v}_i$ & linear velocity of  center of mass for body i   \\ \hline
$\vec{\omega}_i$ & angular velocity of center of mass for body i  \\ \hline
$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
$M$ & generalized mass matrix $ M \in \mathbb{R}^{6n \times 6n}$ (\ref{eq:matrixM})\\ \hline
$I_i$ & inertia tensor for body i  \\ \hline
$C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$  (\ref{eq:matrixN}) \\ \hline
\end {tabular}
\end{center}
\caption {Nomenclature for equations of motion} \label{tab:eom}
\end {table}

\begin{equation} \label{eq:vectorqi}
q_i = \lbrack s_i, x_i, y_i, z_i \rbrack ^T
\end{equation}

\begin{equation} \label{eq:matrixQi} % 4.13 p 50
Q_i = \frac{1}{2} \left[ \begin{array}{ccc}
-x_i & -y_i & -z_i \\
s_i & z_i & -y_i \\
-z_i & s_i & x_i \\
y_i & -x_i & s_i
\end{array} \right]
\end{equation}

\begin{equation} \label{eq:matrixS}
S =  \left[ \begin{array}{ccccc}
1 &  &  & & 0 \\
 & Q_i  \\
 & & \ddots  \\
 & & & 1 \\
0 & & & & Q_n 
\end{array} \right]
\end{equation}

\begin{equation} \label{eq:matrixM}
M =  \left[ \begin{array}{ccccc}
m_i 1 &  &  & & 0 \\
 & I_1  \\
 & & \ddots  \\
 & & & m_n 1 \\
0 & & & & I_n 
\end{array} \right]
\end{equation}

\begin{equation} \label{eq:matrixN}
N =  \left[ \begin{array}{cccc}
\vec{n}_1 &  & & 0 \\
 & \vec{n}_2  \\
 & & \ddots  \\
0  & & & \vec{n}_k 
\end{array} \right]
\end{equation}

@@

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (O1) at(1,1);
\coordinate (O2) at(1,3);
\coordinate (C) at(1,2);
\draw (0,0) -- (2,0) -- (2,2) -- (0,2) --(0,0);
\draw (0,2) -- (2,2) -- (2,4) -- (0,4) --(0,2) ;
\node at (0.5,0.4) {$B_1$};
\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
\node at (0.5,3.5) {$B_2$};
\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{anc}^{\,1}$};
\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{anc}^{\,2}$};
\node[anchor=west] at (4.5,3.5) {
$\vec{r}_i$  =  orientation of body i 
};
\node[anchor=west] at (4.5,3) {$\vec{r}_{anc}^{\,i} $ = anchor point for body i};
\end{tikzpicture}
\caption{Illustration of nomenclature for equations of motion for joint.}
\label{fig:eom-joint}
\end{figure}

@@

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (S) at (2.5,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1.5,0) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma<f_y$}};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}
\coordinate (S) at (3,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1,0) -- (1,2.5) -- (3,2.5) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma=f_y$}};
\end{tikzpicture}
\caption{Stress distribution under elastic and plastic loads.}
\label{fig:wp}
\end{figure}

For rectangular section lagrange multiplier limits
in equations \ref{eq:lambdaLow} and  
\ref{eq:lambdaHigh}
i.e. maximum forces 
are summarized in figure \ref{fig:maxForces}.
Yield shear stress is assumed here to be $ 0.5\, f_y$ using Tresca yield critetion.
If von Mises yield criterion is used 0.5 is replaced by 0.58 ($1/\sqrt{3}$).
These are not exact values in multiaxial stress state but they
should be quite usable in most cases.
For rotation around y-axis there is closed form solution only for
circular cross sections.
Given approximation is 
best suited for cases where b and h are quite similar.
Better approximation for any given b and h can be obtained 
by integrating distance from center of joint over cross section and
multiplying it with yield shear stress e.g. using octave.

\begin{lstlisting}
b=0.01
h=0.01
fy=200e6
wpy=fy/2*dblquad(@(x,y) sqrt(x.*x+y.*y),-b/2,b/2,-h/2,h/2)
38.2
\end{lstlisting}

\lstset{language=octave}
\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (O) at (0,0,0);

% axes
\draw[->] (O) -- (1,0,0) node[anchor=north east]{$x$};
\draw[->] (O) -- (0,1,0) node[anchor=north west]{$y$};
\draw[->] (O) -- (0,0,1) node[anchor=south]{$z$};

@@

\node[anchor=west] at (4.0,3) {
\begin{tabular}{l}
maximum force in x direction = $0.5\, b\, h f_y$ \\
maximum force in y direction = $b\, h\, f_y$  \\
maximum force in z direction = $0.5\, b\, h\, f_y$  \\
\\
maximum moment around x-axis = $0.25\, b\, h^2 \, f_y$  \\
maximum moment around y-axis $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\
maximum moment around z-axis = $0.25\, b^2\, h\, f_y$ 
\end{tabular}
};

>>>>

% pdf page 33, notation in typical ODEs
\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$\vec{r}_i$ & position of center of mass for body i  \\ \hline
$\vec{q}_i$ & orientation for body i as quaternion $\lbrack s_i, x_i, y_i, z_i \rbrack ^T $ \\
\hline
$\vec{p}_i$ & contact or joint point i  \\ \hline
$\vec{r}_{ki}$ & $\vec{p}_k - \vec{r}_i$  \\ \hline
$\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
$Q_i$ & rotation matrix  $
\frac{1}{2} \left[ \begin{array}{ccc}
-x_i & -y_i & -z_i \\
s_i & z_i & -y_i \\
-z_i & s_i & x_i \\
y_i & -x_i & s_i
\end{array} \right]
$
\\ \hline
 & generalized transformation matrix $ S \in \mathbb{R}^{7n \times 6n}$ \\ 
$S$ & $
  \left[ \begin{array}{ccccc}
1 &  &  & & 0 \\
 & Q_i  \\
 & & \ddots  \\
 & & & 1 \\
0 & & & & Q_n 
\end{array} \right]
$
\\ \hline
$\vec{v}_i$ & linear velocity of  center of mass for body i   \\ \hline
$\vec{\omega}_i$ & angular velocity of center of mass for body i  \\ \hline
$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
 & generalized mass matrix $ M \in \mathbb{R}^{6n \times 6n}$ \\ 
$M$ &
$
\left[ \begin{array}{ccccc}
m_i 1 &  &  & & 0 \\
 & I_1  \\
 & & \ddots  \\
 & & & m_n 1 \\
0 & & & & I_n 
\end{array} \right]
$
\\ \hline
$I_i$ & inertia tensor for body i  \\ \hline
$C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$ \\ \hline
\end {tabular}
\end{center}
\caption {Nomenclature for equations of motion} \label{tab:eom}
\end {table}

@@

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (O1) at(1,1);
\coordinate (O2) at(1,3);
\coordinate (C) at(1,2);
\draw (0,0) -- (2,0) -- (2,2) -- (0,2) --(0,0);
\draw (0,2) -- (2,2) -- (2,4) -- (0,4) --(0,2) ;
\node at (0.5,0.4) {$B_1$};
\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
\node at (0.5,3.5) {$B_2$};
\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{anc}^{\,1}$};
\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{anc}^{\,2}$};
\node[anchor=west] at (4.5,3.5) {
$\vec{r}_i$  =  orientation of body i 
};
\node[anchor=west] at (4.5,3) {$\vec{r}_{anc}^{\,i} $ = body frame vector i};
\end{tikzpicture}
\caption{Illustration of nomenclature for equations of motion for joint.}
\label{fig:eom-joint}
\end{figure}

@@

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (S) at (2.5,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1.5,0) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma<f_y$}};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}
\coordinate (S) at (3,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1,0) -- (1,2.5) -- (3,2.5) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma=f_y$}};
\end{tikzpicture}
\caption{Stress distribution for bending under elastic and plastic loads.}
\label{fig:wp}
\end{figure}

Plasticity is handled by defining maximum forces
in equations \ref{eq:lambdaLow} and  
\ref{eq:lambdaHigh} using plastic capasities as maximum forces.

Maximum force acting in same direction as $\vec{r}_{anc}^{\,i} $
is product of area and yield stress as given in equation \ref{eq:fN}.

\begin{equation} \label{eq:fN}
N_{max}= \int_A f_y
\end{equation}

Maximum forces acting perpendicular to $\vec{r}_{anc}^{\,i} $
are product of area and shear yield stress $\tau_y$ as given in equation \ref{eq:fQ}.
\begin{equation} \label{eq:fQ}
Q_{max}= \int_A \tau_y
\end{equation}

Maximum moments acting around axis perpendicular to $\vec{r}_{anc}^{\,i} $
are integrals of perpdendicular distance 
and yield stress $f_y$ as given in equations \ref{eq:Mx} for moment around
x-axis and \ref{eq:Mz} for moment around z-axis.

\begin{equation} \label{eq:Mx}
M_{max}^x= \int_A z f_y
\end{equation}

\begin{equation} \label{eq:Mz}
M_{max}^z= \int_A x f_y
\end{equation}

Maximum moment around $\vec{r}_{anc}^{\,i} $
is integral of distance $d$ from joint point
and shear yield stress $\tau_y$ as given in equations \ref{eq:My}.

\begin{equation} \label{eq:My}
M_{max}^y= \int_A d \tau_y
\end{equation}

Maximum forces and moments for 
rectangular section using constant yield stress
are summarized in figure \ref{fig:maxForces}.
Yield shear stress is assumed here to be $ 0.5\, f_y$ using Tresca yield critetion.
If von Mises yield criterion is used 0.5 is replaced by 0.58 ($1/\sqrt{3}$).
These are not exact values in multiaxial stress state but they
should be quite usable in most cases.

For rotation around y-axis there is closed form solution only for
circular cross sections.
Given approximation is 
best suited for cases where b and h are quite similar.
Better approximation for any given b and h can be obtained 
by integrating distance from center of joint over cross section and
multiplying it with yield shear stress e.g. using octave.

\lstset{language=octave}
\begin{lstlisting}
b=0.01; h=0.01; fy=200e6;
wpy=fy/2*dblquad(@(x,y) sqrt(x.*x+y.*y),-b/2,b/2,-h/2,h/2)
38.2
\end{lstlisting}

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (O) at (2,2,0);

% axes
\draw[->] (O) -- ++(1,0,0) node[anchor=north east]{$x$};
\draw[->] (O) -- ++(0,1,0) node[anchor=north west]{$y$};
\draw[->] (O) -- ++(0,0,1) node[anchor=south]{$z$};

@@

\node[anchor=west] at (4.0,3) {
\begin{tabular}{l}
maximum force in x and z directions = $0.5\, b\, h f_y$ \\
maximum force in y direction = $b\, h\, f_y$  \\
\\
maximum moment around x-axis = $0.25\, b\, h^2 \, f_y$  \\
maximum moment around y-axis $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\
maximum moment around z-axis = $0.25\, b^2\, h\, f_y$ 
\end{tabular}
};
